Master thesis

### 2017-06-12 docs/master-thesis.tex
A4. Sensitivity analysis

@@

>>>>

To study resistance and resilience of ecosystem face to disturbance,
highlighting the role of biodiversity, we decided to use TROLL model
simulations \citep{Chave1999}. In order to get a finer study of
simulations response, we need to assess first sensitivity of the TROLL
model to different parameters. More particularly we need to assess the
importance of functional traits to further better control functional
diversities in simulations. We also need to assess sensitivity of the
model to see rain constant because we assume it is one of the main
factor of tree recruitments after disturbance in the model.

\subsection{Material and methods}\label{material-and-methods-2}

To assess the sensitivity of TROLL model to species functionnal traits,
we performed a sensitivity analysis by fixing species trait values to
their mean. Each trait was tested independently. We reduce to a common
mean traits with a correlation \(r \geq 0.8\) (see figure
\ref{fig:A4corr}).

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/A4corr-1.pdf}
\caption{\label{fig:A4corr}Correlation of functional traits within TROLL
model species Blue represents negative coorelations whereas red
represents positive correlations. Values and colour intensity represents
correlation values.}
\end{figure}

To assess the sensitivity of TROLL model to seed rain, we performed a
sensitivity analysis by fixing simulations seed rain constant to 2, 20,
200 and 2000 seeds per hectare.

Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and
188.9 GB of memory. We assumed maturity of the forest after 500 years of
regeneration \citep{Li} and computed simulation 100 years after a
disturbance event of 40\% intensity. Due to computer limitations we did
not run replicate (besides it should be necessary to reduce simulation
stochasticity). To assess ecosystem outputs sensitivity to studied
parameters, we compared it to 100 replicates of control simulations with
all parameters set to default values. Ecosystem outputs outside of the
range of the control replicates values are significantly influenced by
the studied parameter.

\subsection{Results}\label{results-2}

\subsubsection{Control}\label{control}

Both disturbed ecosystem structure and functional composition
corresponded to ecosystem structure and functional composition before
disturbance (figure \ref{fig:A4controlStr}). Consequently, we assumed
that disturbance did not affect much ecosystem structure and function.
Secondly range of values inside control replicates is low (figure
\ref{fig:A4controlVar}).

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/A4controlStr-1.pdf}
\caption{\label{fig:A4controlStr}Ecosystem structure before disturbance and
disturbed. Ecosystem structure before disturbance (left) and disturbed
(right) with diameter structure (A, B), diversity at different orders
(C, D) and rank-abundance diagrams (E, F).}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/A4controlVar-1.pdf}
\caption{\label{fig:A4controlVar}Control replicates variation. Maximum, mean
and minimum number of trees with dbh above 10 cm (N10) and 30 cm (N30),
above ground biomass (AGB) and basal area (BA) over simulation time.}
\end{figure}

\subsubsection{Functional traits}\label{functional-traits}

Most of functional traits had a significant long term influence on
ecosystem outputs (figure \ref{fig:A4ftVar}). Only \textbf{seed volume}
was always in the range of variation of control replicates. On the other
hand, few functional traits influenced final ecosystem structure (figure
\ref{fig:A4ftStr}). Only specific maximum diameter \textbf{dmax} add
higher diversity for greater orders implying better eveness in species
distributions. Regarding functional composition, traits fixed to mean
did not change other functional traits density distribution.

\textbf{ah-hmax} traits fixed to mean increased number of stems above 10
and above 30 cm dbh and basal area after disturbance (but not in long
term) and did not affect aboveground biomass. Similarly, wood specific
gravity \textbf{wsg} trait fixed to mean had exactly the same effect on
number of stems above 10 and above 30 cm dbh and basal area after
disturbance than \textbf{ah-hmax} but with a time lag ; and \textbf{wsg}
also increased aboveground biomass. \textbf{dmax} trait fixed to mean
slightly decreased number of stems above 10 and above 30 cm dbh over
time while it increased basal area, aboveground biomass, and species
eveness. Finally, leaf mass per area \textbf{LMA} trait fixed to mean
only decreased number of stem above 10 cm dbh after ecosystem resilience
to disturbance (approximately 50 years) but did not affect other
ecosystem outputs.

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/A4ftVar-1.pdf}
\caption{\label{fig:A4ftVar}Functional traits effect on simulation ecosystem
variations over time. Number of trees with dbh above 10 cm (N10) and 30
cm (N30), above ground biomass (AGB) and basal area (BA). Grey area
represents the interval of control replicates whereas black line
represents the mean of control replicates.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/A4ftStr-1.pdf}
\caption{\label{fig:A4ftStr}Functional traits effect on simulation ecosystem
final structure. Tree final height histogram for traits (A) and control
(B), tree final diameter histogram for traits (C) and control (D),
ecosystem final diversity plot at different orders (E), and ecosystem
final rank-abundance diagram (F).}
\end{figure}

\subsubsection{Seed rain}\label{seed-rain}

Seedrain constant affected ecosystem outputs only when set lower than
default value (figure \ref{fig:A4srVar} \& \ref{fig:A4srStr}). Moreover,
seedrain did not seem to affect aboveground biomass (figure
\ref{fig:A4srVar}) and final ecosystem height and diameter structure
(figure \ref{fig:A4srStr}). Seedrain constant fixed to 2 or 20 seed per
hectare seemed to have a similar effect. Lower seedrain implied faster
decrease of stem above 10 cm dbh and higher number of stem above 30 cm
dbh after ecosystem resilience to disturbance (approximately 50 years).
Lower seedrain than default decreased basal area over time. In addition,
lower seedrain than default decreased equitability by increasing
abundance of abundant species and decreasing abundance of less abundant
species. Seedrain constant even decreased the total number of species
when fixed to 2 seed per hectare. Finally, seedrain constant slightly
affected functional composition with higher pike on ecosystem most
representatives functional trait values. In a nutshell, the lower is the
seedrain constant the most the functional density distribution is
aggregated around few functional trait values.

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/A4srVar-1.pdf}
\caption{\label{fig:A4srVar}Seed rain effect on simulation ecosystem
variations over time. Number of trees with dbh above 10 cm (N10) and 30
cm (N30), above ground biomass (AGB) and basal area (BA). Grey area
represents the interval of control replicates whereas black line
represents the mean of control replicates.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/A4srStr-1.pdf}
\caption{\label{fig:A4srStr}Seed rain effect on simulation ecosystem final
structure. Tree final height histogram for traits (A) and control (B),
tree final diameter histogram for traits (C) and control (D).}
\end{figure}

\subsection{Discussion}\label{discussion-1}

\subsubsection{Disturbance simulation}\label{disturbance-simulation}

Ecosystem structure, species organisation and functional composition
stayed the same before and after disturbance. We can thus validate
disturbance module in its actual state. Morevoer we can now consider
that the composition we will initialise at the beginning of simulations
will stay the same after disturbance. Finally control replicates has
shown few stochasticity, it advocates for few or no replicates in
further analysis.

\subsubsection{Functional traits
selection}\label{functional-traits-selection}

\textbf{ah-hmax} fixed to mean implied no high or low trees. Less high
trees left space for more trees increasing number of stems in the
ecosystem thus increasing basal area. Wood specific gravity \textbf{wsg}
fixed to mean mainly increased wood density of light wood species.
Globally higher wood density increased lifespan of individuals
responsible for the time lag and the higher number of stems increasing
basal area. \textbf{wsg} fiwed to mean also increased carbon capture by
individuals, thus increasing aboveground biomass. Specific maximum
diameter \textbf{dmax} fixed to mean decreased death rates. Decreased
death rate diminished number of stems, expecially big ones thus
increasing global basal area and abiveground biomass. \textbf{dmax}
fixed to mean by keeping more small diameter stems also increased random
selection of species increasing eveness in species distribution.
Considering the high correlation between \textbf{ah} and \textbf{hmax}
we could also keep only \textbf{hmax} (because of its more
straightforward ecological meaning).

\subsubsection{Seed rain constant
influence}\label{seed-rain-constant-influence}

Seedrain constant did not directly affect ecosystem global outputs over
simulation post-disturbance time but have a major effect on species and
functional composition and diversity. Reducing seedrain constant
resulted in an ecosystem selecting few species increasing their
abundance and functional dominance of their traits. Thus reduced
seedrain constant greatly diminsihed eveness untill a decrease of total
number of species for its lowest value.

@@

\begin{verbatim}
## 
## Attaching package: 'dplyr'
\end{verbatim}

\begin{verbatim}
## The following objects are masked from 'package:stats':
## 
##     filter, lag
\end{verbatim}

\begin{verbatim}
## The following objects are masked from 'package:base':
## 
##     intersect, setdiff, setequal, union
\end{verbatim}
